---
title: "1-GPS Data processing 2015 - 2 - 240/10"
author: "Aurore Ponchon"
date: "26/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE,dpi=300)
```

GPS loggers are not the same for successful and failed breeders so they have to be processed differently before being merged. Moreover, data from 2015 and 2018 have different formats and different constraints so they are also processed separately. This file aims at formatting the data and compare results for 2015 when cleaning trips based on different threshold parameters.

## Bird data import for 2015
### Import data for successful breerders

Loggers deployed on successful breeders are IgotU GPS. They give __LOCAL TIME__ in UTC (GMT+5) and record locations every 5 minutes.

```{r import of successful breeders 2015}
library(fields)
library(maptools)
library(maps)
library(stringr)
library(xlsx)
library(plotrix)
library(pastecs)
library(lme4)
library(simr)
library(kableExtra)
library(plyr)
source("R scripts/Functions.R")

#Coordinates of the colony
colony<-data.frame(Longitude =77.52356,Latitude=-37.8545)

##Import data for successful birds (classical GPS)
success.2015<-read.table("RawData/ams_yna_2015-16_gps_success.txt",header=T,sep="\t")
success.2015$Status<-as.factor("success")
success.2015$DateTime<-as.POSIXct(paste(strptime(success.2015$Date,format="%d/%m/%Y"),
                                        success.2015$Time,sep=" "),tz="")
success.2015<-success.2015[,c("Logger.ID","Status","DateTime","Longitude","Latitude","Speed","Altitude")]

#distance between location and colony
success.2015$Distmax<-as.vector(rdist.earth(success.2015[,c("Longitude","Latitude")],
                                       colony[1,c("Longitude","Latitude")],miles=F))  
names(success.2015)[1]<-"ID"
success.2015$Distadj<-0
success.2015$Difftime<-0

#calculate distance between 2 consecutive points and time elapsed between 2 consecutive points
for ( n in 2:nrow(success.2015)){
  success.2015$Distadj[n]<-rdist.earth(success.2015[n,c("Longitude","Latitude")],
                                  success.2015[n-1,c("Longitude","Latitude")],miles=F)
   success.2015$Difftime[n]<-difftime(success.2015$DateTime[n],success.2015$DateTime[n-1],units="mins")
}

#remove time difference when changing individuals
success.2015$Difftime[which(success.2015$Difftime<0)]<-0
```

### Import data for failed breeders

Loggers deployed on failed breeders are GPS-UHF Ecotone. They give __time in GMT__ so 5h have to be added to obtain local time (UTC=GMT+5).They record locations every 5-15 minutes, depending on weather conditions. They provide more gaps in the data

```{r import failed breeders 2015}

##Import data for failed birds (GPS-UHF)
fail.2015<-read.table("RawData/ams_yna_2015-16_gps_fail.txt",header=T,sep="\t")
fail.2015$Date<-paste(fail.2015$Year,fail.2015$Month,fail.2015$Day,sep="/")
fail.2015$Time<-paste(fail.2015$Hour,fail.2015$Minute,fail.2015$Second,sep=":")
fail.2015$Status<-as.factor("fail")
fail.2015$DateTime<-as.POSIXct(paste(strptime(fail.2015$Date,format="%Y/%m/%d"),fail.2015$Time,sep=" "),tz="UTC") 
fail.2015$DateTime<-fail.2015$DateTime + 5*3600
fail.2015<-fail.2015[,c("Logger.ID","Status","DateTime","Longitude","Latitude","Speed","Altitude","In.range")]

#when "in.range" is true, locations are set to colony coordinates
fail.2015$Longitude[which(is.na(fail.2015$Longitude)==T & fail.2015$In.range==1)]<-colony$Longitude   
fail.2015$Latitude[which(is.na(fail.2015$Latitude)==T & fail.2015$In.range==1)]<-colony$Latitude 
fail.2015<-fail.2015[!is.na(fail.2015$Latitude)==T,]

#distance between location and colony
fail.2015$Distmax<-NA
fail.2015$Distmax<-as.vector(rdist.earth(fail.2015[,c("Longitude","Latitude")],
                                         colony[1,c("Longitude","Latitude")],miles=F))
fail.2015$Distmax[fail.2015$In.range==1]<-0
names(fail.2015)[1]<-"ID"
fail.2015<-fail.2015[,-8]
fail.2015$Distadj<-0

#calculate distance between 2 consecutive points and time elapsed between 2 consecutive points
fail.2015$Difftime<-0
for (n in 2:nrow(fail.2015)){
  fail.2015$Distadj[n]<-rdist.earth(fail.2015[n,c("Longitude","Latitude")],
                               fail.2015[n-1,c("Longitude","Latitude")],miles=F)
  fail.2015$Difftime[n]<-difftime(fail.2015$DateTime[n],fail.2015$DateTime[n-1],units="mins")
}
fail.2015$Difftime[which(fail.2015$Difftime<0)]<-0

```

```{r plot of birds 2015 all,echo=F,fig.width=15,fig.height=6}
par(mfrow=c(1,2))
plot(success.2015$DateTime,success.2015$Distmax,xlab="Date",ylab="Distance to the colony (km)",
     col=tim.colors(length(unique(success.2015$ID)))[as.factor(success.2015$ID)],pch=16,
     main="Distance to the colony success 2015")
plot(success.2015$DateTime,success.2015$Difftime/60,xlab="Date",ylab="Time difference (hrs)",
     col=tim.colors(length(unique(success.2015$ID)))[as.factor(success.2015$ID)],pch=16,
     main="Time elapsed success 2015")


par(mfrow=c(1,2))
plot(fail.2015$DateTime,fail.2015$Distmax,xlab="Date",ylab="Distance to the colony (km)",
     col=tim.colors(length(unique(fail.2015$ID)))[as.factor(fail.2015$ID)],pch=16,
     main="Distance failed 2015")
plot(fail.2015$DateTime,fail.2015$Difftime/60,xlab="Date",ylab="Time difference (hrs)",
     col=tim.colors(length(unique(fail.2015$ID)))[as.factor(fail.2015$ID)],pch=16,
     main="Time elapsed failed 2015")

```

### Merge data 2015, define trips and clean trips.

Parameters to define a trip are based on distance to the colony, number of locations out of the colony and duration. They are different for data from 2015 to account for duty cycles in loggers tracking successful breeders in 2018.

```{r define and clean trips 2015}
dataset.2015<-rbind(success.2015,fail.2015)
dataset.2015<-dataset.2015[duplicated(dataset.2015$DateTime)==FALSE,] # remove duplicated points with the same date and time
datasetTravelNb<-"NA"
dataset.2015$PathLength<-0

dist.thres<-1             #distance threshold (in km)
last.dist<-10           #distance threshold (in km) when having only one location between 2 trips
row.thres<- 12               #number of rows constituting a trip (corresponding to minimum 1h)
dur.thres<-120         #min duration of a trip
diff.thres<-4*60     #max difftime in a trip
date.max<-as.POSIXct("2015-12-22 12:00:00", format="%F %H:%M:%S",tz="UTC") 


#Define trip number within each individual
ALL.2015<-define_trips(dataset.2015)   #function from source file

#summarize raw trips by individuals or status
dist.max.all.trips.2015.ind<-raw_trips_summary_ind(ALL.2015)    #function from source file
dist.max.all.trips.2015.status<-raw_trips_summary_status(ALL.2015)  #function from source file

#clean trips and give summary by individuals or status
summary.clean.trips.2015.ind<-clean_trips_summary_ind(ALL.2015)    #function from source file
summary.clean.trips.2015.status<-clean_trips_summary_status(ALL.2015)    #function from source file


#clean trips and give locations of selected trips
clean.trips.loc.2015<-clean_trips_locations(ALL.2015)
```

### Summary of raw trips by status

The first table shows the summary of trip characteristics with raw data. The second one shows the summary of trip characteristics with a minimum distance of `r dist.thres`km, a minimum duration of `r dur.thres`, gaps of less than `r diff.thres` and a minimum of `r row.thres` number of locations to define a trip.


```{r, echo=F}
kbl(dist.max.all.trips.2015.status) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


kbl(summary.clean.trips.2015.status) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

### Plots of trips by individuals

In the plots, all black dots represent trips/data that have been excluded. All colored trips are the selected ones.

```{r plot of trips 2015,echo=F,fig.heigth=8,fig.width=15}

levo<-unique(clean.trips.loc.2015$ID)

for (i in 1:length(levo)){
    par(mfrow=c(1,3))
  
    sub<-subset(ALL.2015,ALL.2015$ID==levo[i])
    subo<- subset(clean.trips.loc.2015,clean.trips.loc.2015$ID==levo[i])
    sub$TravelNb<-subo$TravelNb[match(sub$DateTime,subo$DateTime)]
    sub$TravelNb[which(is.na(sub$TravelNb)==T)]<-0
    
    plot(sub$DateTime,sub$Distmax,xlab="Date",ylab="Distance to the colony (km)",
     col=tim.colors(length(unique(ALL.2015$ID)))[i],pch=16,
     main=paste(levo[i],sub$Status[1], sep=" "))

  plot( sub$DateTime,sub$Difftime/60,xlab="Date",ylab="Time elapsed between 2 locations (hrs)",
     col=tim.colors(length(unique(ALL.2015$ID)))[i],pch=16,
     main=paste(levo[i],sub$Status[1], sep=" "))

  plot(sub$DateTime,sub$Distmax,xlab="Date",ylab="Distance to the colony (km)",
     col=ifelse(sub$TravelNb>0,tim.colors(length(unique(ALL.2015$ID)))[i],"black"),pch=16,
     cex=ifelse(sub$TravelNb>0,1.2,0.8),
     main=paste(levo[i],subo$Status[1], sep=" "))
}

```

### Interpolation of clean trips

```{r interpolation 2015}
 reso <- 10 * 60       #time resolution between 2 points in seconds

new.trip<-NULL
loc.interpolated.2015<-NULL
id<-unique(clean.trips.loc.2015$ID)

 #interpolate data for each individual separately 
for (x in 1:length(id)){
  sub<-subset(clean.trips.loc.2015,clean.trips.loc.2015$ID==id[x])
  trip<-unique(sub$TravelNb)

   #interpolate data for each trip separately 
  for (y in 1:length(trip)){
    sub.trip<-subset(sub,sub$TravelNb==trip[y])
    sub.trip$TimeSinceOrigin<-rep(0,nrow(sub.trip))

  for (z in 1:nrow(sub.trip)){
    sub.trip$TimeSinceOrigin[z]<-difftime(sub.trip$DateTime[z],sub.trip$DateTime[1],units="sec")
  }

 
  # Resampling of long and lat with regular time intervals 
  sub.lat1 <- regul(x=sub.trip$TimeSinceOrigin, y=sub.trip$Latitude,
                    n=round((max(sub.trip$TimeSinceOrigin)/reso),0),
                    deltat=reso, methods="linear",xmin=0,units="sec")
  sub.lon1 <- regul(x=sub.trip$TimeSinceOrigin, y=sub.trip$Longitude,
                    n=round((max(sub.trip$TimeSinceOrigin)/reso),0),
                    deltat=reso, methods="linear",xmin=0,units="sec")

  new.sub <- data.frame(Longitude=sub.lon1[[2]]$Series, Latitude=sub.lat1[[2]]$Series, Time=as.vector(sub.lon1[[1]]),
                        DateTime=sub.trip$DateTime[1]+as.vector(sub.lon1[[1]]),TravelNb=trip[y])

  #Recreate a dataframe with individual and trip characteristics
  new.sub$ID<-id[x]
  new.sub$Status<-rep(unique(sub$Status),nrow(new.sub))
  new.sub$Distmax<-as.vector(rdist.earth(new.sub[,c("Longitude","Latitude")],
                                         colony[1,c(1,2)],miles=F))
  new.sub$Difftime<-c(0,difftime(new.sub$DateTime[2:nrow(new.sub)],new.sub$DateTime[1:nrow(new.sub)-1],units="mins"))
  new.sub$Distadj<-0

  for (z in 2:nrow(new.sub)){
    new.sub$Distadj[z]<-rdist.earth(new.sub[z,c("Longitude","Latitude")],
                                    new.sub[z-1,c("Longitude","Latitude")], miles=F)
  }

 new.trip<-rbind(new.trip,new.sub)
  }
}

#create a new global dataset with interpolated data
loc.interpolated.2015<-new.trip[,c("ID","Status","TravelNb","DateTime","Longitude","Latitude",
                                   "Distmax","Difftime","Distadj")]

#save(loc.interpolated.2015,file="NewlyCreatedData/clean_interp_loc_2015.RData")


summary.trips.interp1<-ddply(loc.interpolated.2015,.(ID,Status,TravelNb),summarize,DistmaxIndTrip=max(Distmax),
                             TripDurIndTrip=sum(Difftime)/60,TotPathIndTrip=sum(Distadj))

summary.trips.interp2<-ddply(summary.trips.interp1,.(ID,Status),summarize,NbTrips=length(DistmaxIndTrip),
                             DistmaxInd=mean(DistmaxIndTrip),
                             TripDurInd=mean(TripDurIndTrip),TotPathInd=mean(TotPathIndTrip))

summary.trips.interp<-ddply(summary.trips.interp2,.(Status),summarize,
                            NbTrip=mean(NbTrips),NbTripSE=std.error(NbTrips,na.rm=T),
                            Distmax=mean(DistmaxInd),DistmaxSE=std.error(DistmaxInd,na.rm=T),
                            TripDur=mean(TripDurInd),TripDurSE=std.error(TripDurInd,na.rm=T),
                            TotPath=mean(TotPathInd),TotPathSE=std.error(TotPathInd,na.rm=T))
```
### Summary of clean interpolated trips by status

The first table shows the summary of trip characteristics with raw data. The second one shows the summary of trip characteristics with a minimum distance of `r dist.thres`km, a minimum duration of `r dur.thres`, gaps of less than `r diff.thres` and a minimum of `r row.thres` number of locations to define a trip with a time resolution of `r reso` for interpolation.

```{r, echo=FALSE}
kbl(summary.trips.interp) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Statistics

Linear mixed models are applied to test whether the trip variables are different between failed and successful breeders

```{r LMM, message=F}
maxdist<-lmer(sqrt(DistmaxIndTrip) ~ Status + (1|ID),data=summary.trips.interp1)
dur<-lmer(sqrt(TripDurIndTrip) ~ Status + (1|ID),data=summary.trips.interp1)
totdist<-lmer(sqrt(TotPathIndTrip) ~ Status + (1|ID),data=summary.trips.interp1)

anova(maxdist)
anova(dur)
anova(totdist)
```

### Power analysis based on Simr package

```{r SIMR, echo=FALSE,eval=F,message=FALSE}
simmaxdist<-powerSim(maxdist)
simdur<-powerSim(dur)
simdtotdist<-powerSim(totdist)
```

```{r SIMR2, echo=FALSE,message=FALSE}
summary(simmaxdist)
summary(simdur)
summary(simdtotdist)
```


```




